Data Exploration

### 2019-01-23 src/data_exploration.Rmd
additional vif

>>>>

As per SLR suggestion, rerun without gear

```{r}
vif_allbutgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutgear, "../output/bio/vif_allbutgear.csv", row.names = TRUE)
vif_allbutgear
```

As per SLR suggestion, rerun without gear and most highly correlated variables

```{r}
vif_allbutgearhighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbutgearhighlycorr, "../output/bio/vif_allbutgearhighlycorr.csv", row.names = TRUE)
vif_allbutgearhighlycorr
```

As per SLR suggestion, rerun with gear but without most highly correlated variables

```{r}
vif_allbuthighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter + gear, data = presab))
write.csv(vif_allbuthighlycorr, "../output/bio/vif_allbuthighlycorr.csv", row.names = TRUE)
vif_allbuthighlycorr

vif_allbuthighlycorr_sq <- read.csv("../output/bio/vif_allbuthighlycorr.csv", header = TRUE)
vif_allbuthighlycorr_sq$GVIF2Dfsq <- vif_allbuthighlycorr_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuthighlycorr_sq, "../output/bio/vif_allbuthighlycorr_sq.csv", row.names = TRUE)
vif_allbuthighlycorr_sq
```

ok remove one variable at a time - leave gear in

remove amo_prev
```{r}
vif_allbutamo_prev <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbutamo_prev, "../output/bio/vif_allbutamo_prev.csv", row.names = TRUE)
vif_allbutamo_prev

vif_allbutamo_prev_sq <- read.csv("../output/bio/vif_allbutamo_prev.csv", header = TRUE)
vif_allbutamo_prev_sq$GVIF2Dfsq <- vif_allbutamo_prev_sq$GVIF..1..2.Df..^2
write.csv(vif_allbutamo_prev_sq, "../output/bio/vif_allbutamo_prev_sq.csv", row.names = TRUE)
vif_allbutamo_prev_sq
```
and leave gear out

remove amo_prev + gear
```{r}
vif_allbutamo_prevgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbutamo_prevgear, "../output/bio/vif_allbutamo_prevgear.csv", row.names = TRUE)
vif_allbutamo_prevgear
```

remove chl_depth
```{r}
vif_allbutchl_depth <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutchl_depth, "../output/bio/vif_allbutchl_depth.csv", row.names = TRUE)
vif_allbutchl_depth

vif_allbutchl_depth_sq <- read.csv("../output/bio/vif_allbutchl_depth.csv", header = TRUE)
vif_allbutchl_depth_sq$GVIF2Dfsq <- vif_allbutchl_depth_sq$GVIF..1..2.Df..^2
write.csv(vif_allbutchl_depth_sq, "../output/bio/vif_allbutchl_depth_sq.csv", row.names = TRUE)
vif_allbutchl_depth_sq
```

remove chl_depth and gear
```{r}
vif_allbutchl_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutchl_depthgear, "../output/bio/vif_allbutchl_depthgear.csv", row.names = TRUE)
vif_allbutchl_depthgear
```

remove ssh_surface
```{r}
vif_allbutssh_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutssh_surface, "../output/bio/vif_allbutssh_surface.csv", row.names = TRUE)
vif_allbutssh_surface

vif_allbutssh_surface_sq <- read.csv("../output/bio/vif_allbutssh_surface.csv", header = TRUE)
vif_allbutssh_surface_sq$GVIF2Dfsq <- vif_allbutssh_surface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbutssh_surface_sq, "../output/bio/vif_allbutssh_surface_sq.csv", row.names = TRUE)
vif_allbutssh_surface_sq
```

remove ssh_surface & gear
```{r}
vif_allbutssh_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutssh_surfacegear, "../output/bio/vif_allbutssh_surfacegear.csv", row.names = TRUE)
vif_allbutssh_surfacegear
```

remove o2_surface
```{r}
vif_allbuto2_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuto2_surface, "../output/bio/vif_allbuto2_surface.csv", row.names = TRUE)
vif_allbuto2_surface

vif_allbuto2_surface_sq <- read.csv("../output/bio/vif_allbuto2_surface.csv", header = TRUE)
vif_allbuto2_surface_sq$GVIF2Dfsq <- vif_allbuto2_surface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuto2_surface_sq, "../output/bio/vif_allbuto2_surface_sq.csv", row.names = TRUE)
vif_allbuto2_surface_sq
```

remove o2_surface & gear
```{r}
vif_allbuto2_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuto2_surfacegear, "../output/bio/vif_allbuto2_surfacegear.csv", row.names = TRUE)
vif_allbuto2_surfacegear
```

so ther is a question between temp-surface and o2_depth (-0.51923)... let's look at this

with gear
```{r}
vif_allbuttemp_surface <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuttemp_surface, "../output/bio/vif_allbuttemp_surface.csv", row.names = TRUE)
vif_allbuttemp_surface

vif_allbuttemp_surface_sq <- read.csv("../output/bio/vif_allbuttemp_surface.csv", header = TRUE)
vif_allbuttemp_surface_sq$GVIF2Dfsq <- vif_allbuttemp_surface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuttemp_surface_sq, "../output/bio/vif_allbuttemp_surface_sq.csv", row.names = TRUE)
vif_allbuttemp_surface_sq

vif_allo2_depth <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allo2_depth, "../output/bio/vif_allo2_depth.csv", row.names = TRUE)
vif_allo2_depth

vif_allo2_depth_sq <- read.csv("../output/bio/vif_allo2_depth.csv", header = TRUE)
vif_allo2_depth_sq$GVIF2Dfsq <- vif_allo2_depth_sq$GVIF..1..2.Df..^2
write.csv(vif_allo2_depth_sq, "../output/bio/vif_allo2_depth_sq", row.names = TRUE)
vif_allo2_depth_sq
```

without gear
```{r}
vif_allbuttemp_surfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuttemp_surfacegear, "../output/bio/vif_allbuttemp_surfacegear.csv", row.names = TRUE)
vif_allbuttemp_surfacegear

vif_allo2_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allo2_depthgear, "../output/bio/vif_allo2_depthgear.csv", row.names = TRUE)
vif_allo2_depthgear
```

#maxent?

```{r}
library("raster")
library("dismo")
library("rgeos")
```
